Symmetric Circuits for Rank Gates

### 2017-11-28 sections/translating-formulas-to-cicuits.tex
28 Nov 2017

\subsection{Circuit Manipulation}
In this section and the next we develop a number of results that reference
$P$-uniform families of circuits. In order to prove these results we often need
to define polynomial-time algorithms for generating or manipulating circuits. In
this subsection we define a number of algorithms that perform commonly invoked
operations on circuits.

We often include quotes around the names of these algorithm in order to avoid
any confusion. We assume a reasonable encoding of the circuit.

@@

\begin{definition}
  The operation $\op{merge}(C, H)$ takes as input a circuit $C = \langle G,
  \Omega, \Sigma , \Lambda, L \rangle$ and a set of syntactically equivalent
  gates $H = \{h_1, \ldots , h_p\}$ and outputs the circuit $C' = \langle G',
  \Omega', \Sigma', \Lambda', L' \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{h_2, \ldots, h_q\}$,
  \item $\Omega' = \Omega$,
  \item $\Sigma' = \restr{\Sigma}{G'}$,
  \item $\Lambda' = \Lambda$, and
  \item For all $g \in G'$ and $\vec{a} \in Ind(g)$ if $L(g) (\vec{a}) \in H$
    then $L'(g)(\vec{a}) = h_1$ and if $L(g)(\vec{a}) \notin H$ then
    $L'(g)(\vec{a}) = L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

It can be shown that $\op{merge}$ can be computed by an algorithm running in
time polynomial in the size of the input circuit. We note that the input and
output circuits compute the same function. Moreover, if the input circuit is
symmetric (resp. has unique labels) then the output circuit will also be
symmetric (resp. have unique labels).

@@

\begin{definition}
  The operation $\op{merge-all}(C, H)$ takes as input an ordered circuit $C$ and
  a set of gates $H$ with unique labels and outputs another circuit $C'$. Let
  $H_1, \ldots, H_p$ be the equivalence classes of $H$ linearly ordered using
  the order on the circuit but such that for all $i, j \in [p]$ if $i \leq j$
  then the height of the gates in $H_i$ is less than or equal to the height of
  the gates in $H_j$. Let $C'$ be the result of successively executing
  $\op{merge}$ on each partition.
\end{definition}

For an input tuple $(C, H)$ it follows from Lemma \cite{} that the syntactic
equivalence relation on $H$ can be computed in time polynomial in the size of
$C$. Thus it can be shown that the operation $\op{merge-all}$ can be computed in
time polynomial in the size of the input circuit. As in the case of
$\op{merge}$, the operation $\op{merge-all}$ preserves symmetry, unique labels
and the ensures that the input and output circuits compute the same function. If
$C$ is an ordered circuit and $H$ is a set of gates with unique labels, then if
$g$ is a gate in both $C$ and $C' := \op{merge-all}(C, H)$, then if all of the input
gates of $g$ in $C$ are in $H$ then no two distinct input gates of $g$ in $C'$
are syntactically equivalent (but $g$ does not have unique labels in $C'$ as
$L'(g)$ may not be injective).

@@

\begin{definition}
  The operation $\op{replace}(C, g_1, g_2)$ takes as input a circuit $C =
  \langle G, \Omega, \Sigma, \Lambda, L\rangle$ and a pair of gates $(g_1, g_2)$
  in $C$ and outputs a circuit $C' = \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$ such that
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{g_1\}$,
  \item $\Omega'$ is an function from the domain of $\Omega$ to $G'$ such that
    for each $\vec{a} \in \dom(Omega)$ if $\Omega(\vec{a}) = g_1$ then
    $\Omega'(\vec{a}) = g_2$ and if then $\Omega'(\vec{a}) = \Omega(\vec{a})$.
  \item $\Sigma' = \restr{\Sigma}{G'}$
  \item $\Lambda' = \restr{\Lambda}{G'}$
  \item For each $g \in G'$ and each $\vec{a} \in \ind(g)$ if $L(g)(\vec{a}) =
    g_1$ then $L'(g)(\vec{a}) = g_2$ and if not then $L'(g)(\vec{a}) =
    L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

@@

  For each gate $h \in G \setminus \{g\}$ and $\vec{b} \in \ind(h)$, if
  $L(h)(\vec{b}) = g$ then let $L'(h)(\vec{b}) = \land_0$, otherwise let
  $L'(h)(\vec{b}) = L(h)(\vec{b})$.

\end{definition}

@@

\begin{definition}
  The operation $\op{concatenate}(C, \mathcal{C}, \tau_R, R)$, where $C :=
  \langle G, \Omega, \Sigma, \Lambda, L \rangle$ is a $(\mathbb{B},
  \tau)$-circuit, $\tau_R := (R_1, \ldots, R_t)$ is a sequence of relation
  symbols such that $\tau_R \subseteq \tau$, $\mathcal{C} := ( C^1, \ldots,
  C^t)$ is a finite sequence of circuits, where for each $i \in [t]$, $C^i =
  \langle G_i, \Omega_i, \Sigma_i, \Lambda_i, L_i \rangle$ is a $(\mathbb{B}_i,
  \tau_i)$-circuit such that $\tau_i \not\subset \tau_R$ and $C^i$ computes a
  query of arity $\arty(R_i)$, and $R$ is a relation symbol in $(\tau \setminus
  \tau_R) \cup (\bigcup_{1 \leq i \leq t}\tau_i))$. The operation outputs a
  circuit $C' = \langle G', \Omega', \Sigma', \Lambda', L' \rangle$ defined from
  the input as follows. Note that all circuits take in structures of size $n$.

@@

It can be shown that $\op(concatenate)$ can be computed in time polynomial in
the combined size of the input circuits. Moreover, if the input circuits are all
symmetric (resp. have unique labels) then the output circuit will be symmetric
(resp. have unique labels). We sometimes refer to $\op(concatenate)$ $(C^1,
\ldots, C^t)$ into $C^0$ along $(R_1 , \ldots , R_t)$ when invoking
`glue-circuit' on the input $(C^0, \ldots , C^t)$ and $\tau_R = \{R_1, \ldots ,
R_t\}$.

\subsection{Unique Labels and Rigidity}
In contrast with the work of Anderson and Dawar~\cite{}, the circuit families we
use to establish our main result are additionally required to have unique
labels. In this subsection we briefly discuss this inclusion, and present
results suggesting its importance.

It is important to note that our analysis of circuits depends crucially on the
fact that a circuit may be transformed (in polynomial time) into an equivalent
circuit with the property that each element of $\sym_n$ induces a unique
automorphism on the circuit $C_n$. This fact ensures a canonical action of
$\sym_n$ on the circuit $C_n$ and is a crucial requirement for defining needed
theoretical tools, such as supports, and computing various facts about the
circuit in polynomial time, such as the coarsest supporting partition or the
orbit of a gate.

Anderson and Dawar~\cite{AndersonD17} define a condition analogous to the one
define above and prove that it's a sufficient condition for establishing the
uniqueness of the induced automorphism on the circuit and that a circuit can be
transformed into an equivalent rigid circuit in polynomial time.

Although the rigidity condition we present in this paper is similarly defined,
the existence of a polynomial-time computable transformation of a circuit into
an equivalent rigid circuit is not as straight forward. Indeed, it is possible
to show that the problem of determining if a given $(\mathbb{B}, \tau)$-circuit
is rigid is as least as hard as the graph isomorphism problem. This is a direct
consequence of the following result.

\begin{prop}
  There is a a polynomial time reduction from the graph isomorphism problem to
  the problem of deciding if a specified gate in a matrix-circuit is syntacticly
  equivalent to any other gate in the circuit.
  \label{prop:syntactic-graph-iso}
\end{prop}
\begin{proof}
\end{proof}

Another obvious consequence of this result is that if the syntactic equivalence
relation for an input $(\mathbb{B}, \tau)$-circuit can be computed in polynomial
time then the graph isomorphism problem is in $\PT$. It is worth noting that if
we restrict the input to a class of circuits defined over symmetric bases then
we can decide the syntactic equivalence relation in polynomial time.

\begin{prop}
  Let $C_n$ be a circuit with symmetric gates. There is an algorithm that takes
  in such a circuit, runs in time polynomial in the size of the circuit, and
  outputs the syntactic equivalence relation on the gates of $C_n$.
  \label{prop:symmetric-syntactic}
\end{prop}
\begin{proof}
\end{proof}

Moreover, it turns out that being able to compute the syntactic equivalence
relation in polynomial time for some class of circuits $\mathcal{C}$ implies
that the existence of a polynomial-time algorithm for converting circuits in $C$
into equivalent rigid circuit.

\begin{prop}
  Let $\mathcal{C}$ be a set of circuits. Suppose the syntactic equivalence
  relation can be computed in polynomial time for circuits in $\mathcal{C}$. It
  follows that there is an algorithm that runs in polynomial time that takes in
  a circuit $C \in \mathcal{C}$ and outputs a rigid circuit $C'$ such that if
  $C$ is symmetric then $C'$ is symmetric.
  \label{prop:syntactic-equivilence-rigid}
\end{prop}
\begin{proof}
\end{proof}

The following result gives us that the unique labels condition is sufficient.

\begin{prop}
  Let $C_n := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
  \tau)$-circuit with unique labels. There is a deterministic algorithm that
  takes in such a circuit and outputs the syntactic equivalence relation on $G$.
  Moreover, this algorithm runs in time polynomial in the size of the circuit.
  \label{prop:unique-labels-syntactic-equiv}
\end{prop}
\begin{proof}
  We proceed to define a relation $\sim$ by induction on height (i.e. distance
  from an input gate). This relation will be the syntactic equivalence relation
  on $G$. It is easy to check the conditions for syntactic equivalence for gates
  of height 0 as all such gates are input gates and so are either constant gates
  or relational gates. Suppose $g$ and $h$ are gates of height greater than
  zero. We only need to check conditions (i), (iv) and (v) of syntactic
  equivalence. It is easy to check (i) and (v). If either condition is false
  then $g$ and $h$ are not syntacticly equivalent. So suppose $g$ and $h$ are
  two internal gates such that $\Sigma (g) = \Sigma(h)$ and if they are both
  output gates then $\Omega^{-1}(g) = \Omega^{-1}(h)$. It remains to check if
  $L(g)/{\sim}$ and $L(h)/{\sim}$ are isomorphism equivalent. First check if
  $H_h /{\sim} = H_g /{\sim}$. If not, $g$ and $h$ are not syntactically
  equivalent. So suppose $H_h /{\sim} = H_g /{\sim}$. We know that $C_n$ has
  unique labels, and so $L(g)/{\sim}$ and $L(h)/{\sim}$ are injections so are
  isomorphism equivalent if, and only if, $L(g)^{-1}L(h)$ acts like an
  isomorphism. This can be checked by first checking that the function induces a
  sorted permutation on the universe of $g$ (which can be done by iterating over
  the index of $g$). This is sufficient as $L(g)(L(g)^{-1}L(h)) = L(h)$.

@@

% \begin{prop}
%   There is an algorithm that takes in a $(\mathbb{B}, \tau)$-circuit $C$ with
%   unique labels and outputs the syntactic equivalence relation on its gates.
%   This algorithm runs in time polynomial in the size of $C$.
%   \label{prop:unique-labels-syntactic-equiv}
% \end{prop}
% \begin{proof}
%   We build of the relation by induction. Let $g$
% \end{proof}

Putting together Propositions \ref{prop:symmetric-syntactic},
\ref{prop:syntactic-equivilence-rigid}, and
\ref{prop:unique-labels-syntactic-equiv}, we have that there is a
polynomial-time algorithm that takes as input a circuit with unique labels and
outputs an equivalent rigid circuit. In fact, there is an equivalence of sorts
between the unique labels condition and the rigidity condition.

\begin{prop}
  Let $\mathcal{C}$ be a family of circuits. There is an algorithm that runs in
  polynomial-time and takes in a circuit $C \in \mathcal{C}$ and outputs a rigid
  circuit $C'$ such that if $C$ is symmetric then $C'$ is symmetric if, and only
  if, there is an algorithm that runs in polynomial-time and takes in a circuit
  $C \in \mathcal{C}$ and outputs a circuit $C'$ with unique labels such that if
  $C$ is symmetric then $C'$ is symmetric.
\end{prop}

It follows then that for the symmetric circuits defined over bases of symmetric
functions discussed by Anderson and Dawar~\cite{} we may assume the circuits are
rigid and/or have unique labels without a loss of generality. For the more
general circuit discussed in this paper we cannot make such an assumption as
deciding rigidity is harder then graph isomorphism. However, if we restrict
ourselves to families of circuits with unique labels then again we may assume
rigidity without a loss of generality.

Of course, this restriction to circuits with unique labels is only useful if we
can show that formulas in rank logic can be translated into $P$-uniform families
circuits with unique labels. In the next subsection we prove that this is indeed
the case.

\subsection{Translating FPR to Symmetric Matrix Circuits}
Anderson and Dawar~\cite{} define a translation from formulas of $\FPC$ into
equivalent $P$-uniform families of symmetric majority-circuits. More formally,
they show that for $\theta(\vec{x}) \in \FPC[\tau]$ there exists a $P$-uniform
family of symmetric majority-circuits $(C_n)_{n \in \nats}$ such that the query
computed by $(C_n)_{n \in \nats}$ is equal to the query defined by $\theta
(\vec{x})$. In this subsection we develop a similar translation from formulas of
$\FPR$ into $P$-uniform families of symmetric matrix-circuits.

>>>>

\subsection{Algorithms for Circuits}
In order to translate between $P$-uniform families of circuits and formulas of
$\FPR$ we will require that certain important circuit transformations can be
completed in polynomial time. In this subsection we define a number of
operations that perform commonly invoked operations on circuits, and which can
easily be shown to be implementable as algorithms that run in polynomial time.
In each case we assume a reasonable encoding of the circuit.

@@

\begin{definition}
  The operation $\op{merge}(C, H)$ takes as input a circuit $C = \langle G,
  \Omega, \Sigma , \Lambda, L \rangle$ and a set of syntactically equivalent
  gates $H := \{h_1, \ldots , h_p\}$ and outputs the circuit $C' = \langle G',
  \Omega', \Sigma', \Lambda', L' \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{h_2, \ldots, h_q\}$,
  \item $\Omega' = \Omega$,
  \item $\Sigma' = \restr{\Sigma}{G'}$,
  \item $\Lambda' = \Lambda$, and
  \item For all $g \in G'$ and $\vec{a} \in Ind(g)$ if $L(g) (\vec{a}) \in H$
    then $L'(g)(\vec{a}) = h_1$ and if $L(g)(\vec{a}) \notin H$ then
    $L'(g)(\vec{a}) = L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

It can be shown that $\op{merge}$ can be computed by an algorithm running in
time polynomial in the size of the input circuit. We note that the input and
output circuits compute the same function. Moreover, if the input circuit is
symmetric (resp. has unique labels, is transparent) then the output circuit will
also be symmetric (resp. have unique labels, is transparent).

@@

\begin{definition}
  The operation $\op{merge-all}(C, H)$ takes as input an ordered circuit $C$ and
  a set of gates $H$ and outputs another circuit $C'$. Let $H_1, \ldots, H_p$ be
  the syntactic equivalence classes of $H$ linearly ordered using the order on
  the circuit but such that for all $i, j \in [p]$ if $i \leq j$ then the height
  of the gates in $H_i$ is less than or equal to the height of the gates in
  $H_j$. Let $C'$ be the result of successively executing $\op{merge}$ on each
  equivalence class.
\end{definition}

For an input tuple $(C, H)$ if $H$ is a set of gates with unique labels it
follows from Lemma \cite{} that the syntactic equivalence relation on $H$ can be
computed in time polynomial in the size of $C$. Thus, if the set $H$ has unique
labels it can be shown that the operation $\op{merge-all}$ can be computed in
time polynomial in the size of the input circuit. As in the case of
$\op{merge}$, the operation $\op{merge-all}$ preserves symmetry, unique labels
and the ensures that the input and output circuits compute the same function. If
$C$ is an ordered circuit and $H$ is a set of gates with unique labels, then if
$g$ is a gate in both $C$ and $C' := \op{merge-all}(C, H)$, then if all of the
input gates of $g$ in $C$ are in $H$ then no two distinct input gates of $g$ in
$C'$ are syntactically equivalent (but $g$ does not have unique labels in $C'$
as $L'(g)$ may not be injective).

@@

\begin{definition}
  The operation $\op{replace}(C, g_1, g_2)$ takes as input a circuit $C =
  \langle G, \Omega, \Sigma, \Lambda, L\rangle$ and a pair of gates $(g_1, g_2)$
  in $C$ and outputs a circuit $C' = \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$ such that
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{g_1\}$,
  \item $\Omega'$ is an function from the domain of $\Omega$ to $G'$ such that
    for each $\vec{a} \in \dom(\Omega)$ if $\Omega(\vec{a}) = g_1$ then
    $\Omega'(\vec{a}) = g_2$ and if then $\Omega'(\vec{a}) = \Omega(\vec{a})$.
  \item $\Sigma' = \restr{\Sigma}{G'}$
  \item $\Lambda' = \restr{\Lambda}{G'}$
  \item For each $g \in G'$ and each $\vec{a} \in \ind(g)$ if $L(g)(\vec{a}) =
    g_1$ then $L'(g)(\vec{a}) = g_2$ and if not then $L'(g)(\vec{a}) =
    L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

@@

  For each gate $h \in G \setminus \{g\}$ and $\vec{b} \in \ind(h)$, if
  $L(h)(\vec{b}) = g$ then let $L'(h)(\vec{b}) = \land_0$, otherwise let
  $L'(h)(\vec{b}) = L(h)(\vec{b})$.
\end{definition}

@@

\begin{definition}
  The operation $\op{concatenate}(C, \mathcal{C}, \tau_R, R)$, where $C :=
  \langle G, \Omega, \Sigma, \Lambda, L \rangle$ is a $(\mathbb{B},
  \tau)$-circuit, $\tau_R := (R_1, \ldots, R_t)$ is a sequence of relation
  symbols such that $\tau_R \subseteq \tau$, $\mathcal{C} := ( C^1, \ldots,
  C^t)$ is a finite sequence of circuits, where for each $i \in [t]$, $C^i =
  \langle G_i, \Omega_i, \Sigma_i, \Lambda_i, L_i \rangle$ is a $(\mathbb{B}_i,
  \tau_i)$-circuit such that $\tau_i \not\subset \tau_R$ and $C^i$ computes a
  query of arity $\arty(R_i)$, and $R$ is a relation symbol in $(\tau \setminus
  \tau_R) \cup (\bigcup_{1 \leq i \leq t},\tau_i))$. The operation outputs a
  circuit $C' = \langle G', \Omega', \Sigma', \Lambda', L' \rangle$ defined from
  the input as follows. Note that all circuits take in structures of size $n$.

@@

It can be shown that $\op(concatenate)$ can be computed in time polynomial in
the sum of the sizes of the input circuits. Moreover, if the input circuits are
all symmetric then the output circuit will be symmetric. Furthermore, if the
input sequences of circuits and relations have just a single element, then if
the input circuits have unique labels then the output circuit will have unique
labels.

\subsection{Transparency and Unique Extensions}
\label{sec:transparency-and-unique-extensions}
% In contrast with the work of Anderson and Dawar~\cite{AndersonD17}, the
% circuit families we use to establish the main result of this paper are
% additionally required to be transparent. In this subsection we briefly discuss
% this condition, and present results suggesting its necessity.

As mentioned earlier, the unique extensions condition plays a crucial role in
the definition of needed theoretical tools, such as supports, and for proving
the existence of polynomial-time algorithms computing important functions, such
as the coarsest supporting partition or the orbit of a gate (see Proposition
\ref{lem:computing support}).

In this subsection we show that if a circuit is transparent then it may be
transformed in polynomial time into an equivalent circuit with unique labels.
Since we already have that circuits with unique labels have unique extensions
(see Proposition \ref{prop:unique-labels-unique-extensions}), it follows that we
may assume, without a loss of generality, that $P$-uniform families of
transparent circuits have unique labels, and hence unique extensions. We also
discuss relevant special cases and the motivate the explicit requirement that
the $P$-uniform families of circuit referenced in the main result of this paper
consist of transparent circuits.

% Anderson and Dawar~\cite{AndersonD17} introduce the notion of a `rigid'
% circuit, a condition analogous to the rigidity condition defined in this
% paper. The following

% \begin{definition}
%   We say a circuit $C_n$ is que labels it has unique extensions. We finally
%   discuss th\emph{rigid} if it has injective labels and for every pair of
%   gates $(g, g')$ in $C_n$ such that $H_g = H_{g'}$ and $g \equiv g'$.
% \end{definition}

We first show that the syntactic equivalence relation can be efficiently
computed for transparent circuits.

\begin{remark}
  I have introduced the notion of a syntactically transparent class. The reason
  I did this seems to have dissipated, I will probably remove this definition
  and simply the statement of the below proposition.
\end{remark}

\begin{definition}
  We say a class of circuits $\mathcal{C}$ is \emph{syntactically transparent}
  if there is an algorithm that takes in a circuit $C \in \mathcal{C}$ and
  outputs the syntactic equivalence relation on the gates of $C$ and runs in
  time polynomial in the size of the input circuit.
\end{definition}

\begin{prop}
  Any class of transparent circuits is syntactically transparent.
  % Let $C_n := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a
  % $(\mathbb{B},
  % \tau)$-circuit with unique labels. There is a deterministic algorithm that
  % takes in such a circuit and outputs the syntactic equivalence relation on
  % $G$.
  % Moreover, this algorithm runs in time polynomial in the size of the circuit.
  \label{prop:unique-labels-syntactic-equiv}
\end{prop}
\begin{proof} Let $C_n := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a
  transparent $(\mathbb{B}, \tau)$-circuit. We define a relation $\sim$ by
  induction on height (i.e. distance from an input gate). This relation will be
  the syntactic equivalence relation on $G$. It is easy to check the conditions
  for syntactic equivalence for gates of height 0 as all such gates are input
  gates and so are either constant gates or relational gates. Suppose $g$ and
  $h$ are gates of height greater than zero. We only need to check conditions
  (i), (iv) and (v) of syntactic equivalence. It is easy to check (i) and (v).
  If either condition is false then $g$ and $h$ are not syntacticly equivalent.
  So suppose $g$ and $h$ are two internal gates such that $\Sigma (g) =
  \Sigma(h)$ and suppose that if they are both output gates then $\Omega^{-1}(g)
  = \Omega^{-1}(h)$. It remains to check if $L(g)/{\sim}$ and $L(h)/{\sim}$ are
  isomorphism equivalent. First check if $H_h /{\sim} = H_g /{\sim}$. If not,
  $g$ and $h$ are not syntactically equivalent. So suppose $H_h /{\sim} = H_g
  /{\sim}$. If $g$ and $h$ are symmetric gates then $H_h /{\sim} = H_g/{\sim}$
  if, and only if, $L(g)/{\sim}$ and $L(h)/{\sim}$ are isomorphism equivalent,
  so let $g \sim h$. If $g$ and $h$ are not symmetric gates then, since $C_n$ is
  transparent, $g$ and $h$ have unique labels. We thus have that $L(g)/{\sim}$
  and $L(h)/{\sim}$ are isomorphism equivalent if, and only if, $L(g)^{-1}L(h)$
  acts like an isomorphism. This can be checked by first checking that the
  function induces a sorted permutation on the universe of $g$ (which can be
  checked efficiently). This is sufficient as $L(g)(L(g)^{-1}L(h)) = L(h)$.

@@

\begin{prop}
  Let $C_n \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a transparent
  $(\mathbb{B}, \tau)$-circuit. There is an algorithm that that takes in such a
  circuit and outputs a $(\mathbb{B} \cup \mathbb{B}_{\std}, \tau)$-circuit $C'$
  such that $C$ and $C'$ compute the same function, $C'$ has unique labels, and
  if $C$ is symmetric then $C'$ is symmetric. Moreover, this algorithm runs in
  time polynomial in the size of the input circuit.
  \label{prop:transparent-unique}
\end{prop}
\begin{proof}
  Let $C' := \op{make-injective-all}(\op{merge-all} (C, G))$. We have that $C'$
  is symmetric if $C$ is symmetric. Since the syntactic equivalence relation can
  be computed in polynomial time for circuits in $\mathcal{C}$ it follows that
  this operation can be implemented so as to run in time polynomial in the size
  of $C$. Clearly every gate $g$ in $C'$ belongs to a singleton syntactic
  equivalence class.
\end{proof}

Anderson and Dawar~\cite{AndersonD17}, in their study of circuits with symmetric
gates, show that such circuits may be transformed in polynomial time into
equivalent circuits with unique extensions. They do this by showing that a
circuit with symmetric gates may be transformed into an equivalent \emph{rigid}
circuit in polynomial time and that all rigid circuits have unique extensions.
We have similarly shown that transparent circuits can be translated into
equivalent circuits with unique labels in polynomial time and that circuits with
unique labels have unique extensions. Since all circuits with symmetric gates
are transparent, this is a generalisation of their result.

It is natural to ask if the requirement in the main theorem of this paper that
the circuits be transparent is strictly necessary. It may be that, just as for
the case of circuits with symmetric gates, a circuit (not necessarily
transparent) can be transformed in polynomial time into an equivalent
transparent circuit. While we do not address this question exactly, we do show
that the problem of computing the syntactic equivalence relation for a circuit
is at least as hard as the graph isomorphism problem. We also show that that the
related problems of checking if given gate in a circuit has unique labels or has
unique extensions (in a sense we define below) are all also at least as hard as
the graph isomorphism problem. We move on to show that computing other basic
facts about the circuit, such as the orbit of a gate, is also at least as hard
as the graph isomorphism problem.

% ,; address two related decision problems in Proposition \ref{} and Corollaries
% \ref{} and \ref{}. In partic

% As we will show in Proposition \ref{}, formulas of $\FPR$ can be translated
% into. For this transparent circuits. However, the transformation is
% complicated by the requirement that the resultant circuit be transparent.
% However, it is worth asking if the transparency condition is required, i.e. if
% it can be shown that

% \emph{rigid} circuit. They show that all circuits such circuits can be
% transformed into equivalent rigid circuits . They show that if a circuit is
% rigid then it has unique extensions and that every circuit with injective
% labels and symmetric gates can be transformed into an equivalent rigid circuit
% in polynomial time. We have similarly shown that a transparent circuit can be
% transformed into an equivalent one with unique labels in polynomial time, and
% that circuits with unique labels have unique extensions. Moreover, since all
% circuits with symmetric gates and injective labels are transparent, this is a
% strict generalisation of their result.

% and they show that a symmetric gate

% Using the language developed in this paper, a circuit $C_n$ is \emph{rigid} if
% it has injective labels and there are no two gates $g$ and $h$ in $C_n$ such
% that $g$ and $h$ are syntactically equivalent. Since their discussion is
% limited to circuits with symmetric gates and injective labels this is equ

% circuit, a condition analogous to the rigidity condition defined in this
% paper. The following

% It is worth noting that all symmetric circuits are transparent, and

% prove that it's a sufficient condition for establishing the uniqueness of the
% induced automorphism on the circuit. Moreover, they prove that any circuit
% (noting that they limit themselves to a discussion of circuits over a
% symmetric basis) may be transformed into an equivalent rigid circuit in
% polynomial time. As such, they may assume $P$-uniform symmetric circuit
% families are rigid without a loss of generality, and hence be may assume
% without a loss of generality that each permutation in $\sym_n$ uniquely
% extends to an automorphism of the circuit.

% However, in this subsection we show that deciding the rigidity of circuits
% over bases which may include non-symmetric functions is at least as hard as
% the graph isomorphism problem. However, we shoudl that on classes

% Although the rigidity condition we present in this paper is similarly defined,
% the existence of a polynomial-time computable transformation of a circuit into
% an equivalent rigid circuit is not as straight forward. Indeed, it is possible
% to show that the problem of determining if a given $(\mathbb{B},
% \tau)$-circuit is rigid is as least as hard as the graph isomorphism problem.
% This is a direct consequence of the following result.

% \begin{definition}
%   The \emph{syntactic equivalence class problem} is the decision problem for
%   the set of all triples $(C, g_1, g_2)$, where $C$ is a $(\mathbb{B},
%   \tau)$-circuit, and $g_1$ and $g_2$ are two syntactically equivalent gates
%   in $C$.
% \end{definition}

% \begin{definition}
%   The \emph{rigidity problem} is the decision problem for the set of all
%   circuits $C$ where $C$ is a rigid circuit.
% \end{definition}

% \begin{definition}
%   The \emph{unique labels problem} is the decision problem for the set of all
%   circuits $C$ where $C$ has unique labels.
% \end{definition}

% \begin{prop}
%   Let $\tau$ be any non-empty relational vocabulary and let $n, r, p \in
%   \nats$ with $p$ prime. There is a a polynomial time reduction from the graph
%   isomorphism problem to the problem of deciding if a given pair of gates in a
%   given symmetric rank-circuit for structures of size $n$ that (i) has
%   injective labels, (ii) contains no constant gates, and (iii) contains at
%   most two rank gates with bound $r$ and prime $p$, are syntactically
%   equivalent.
%   \label{prop:syntactic-graph-iso}
% \end{prop}
% \begin{proof}
%   We have from \cite{} that there is a polynomial time reduction from the
%   graph isomorphism problem to the bipartite isomorphism problem. The
%   bipartite isomorphism problem is the decision problem for the set of all
%   pairs of partitioned bipartite graphs $B_1 := (U_1, V_1, E_1)$ and $B_2 :=
%   (U_2, V_2, E_2)$ such that there exists a graph isomorphism $\pi : B_1
%   \rightarrow B_2$ with $\restr{\pi}{U_1} = U_2$ and $\restr{\pi}{V_1} = V_2$.

%   Suppose we are given two partitioned bipartite graphs $B_1 := (U_1, V_1,
%   E_1)$ and $B_2 := (U_2, V_2, E_2)$. We assume, without a loss of generality,
%   that there exists $a, b \in \nats$ such that $U_1 = U_2 = [a]$ and $V_1 =
%   V_2 = [b]$.

%   Let $R$ be a relation symbol in $\tau$ and let $k := \arty{R}$. Let $G_{R}
%   := \{g_{R, \vec{c}} : \vec{c} \in [n]^k\}$, $G_{\text{mid}} := \{\land_0,
%   \land_1, \neg_0, \land_{\text{out}} \}$, $G_{\rank} : = \{ g_{\rank, 1},
%   g_{\rank, 2}\}$, and $G_{\land} := \{g_{i, (u,v)} : i \in [2] \, (u,v) \in
%   E_i \}$. Let $C_n = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a
%   $(\{\rank^{1}_2\}, \tau)$-circuit defined as follows. Let $G = G_R \cup
%   G_{\text{mid}} \cup G_{\rank} \cup G_{\land}$ and $\Omega$ be the $0$-ary
%   function $\land_{\text{out}}$. For each $\vec{c} \in [n]^{\arty{R}}$ let
%   $\Lambda(\vec{c}) = g_{R, \vec{c}}$. Define $\Sigma$ as follows. For each $g
%   \in G$,
%   \begin{itemize}
%     \setlength\itemsep{0mm}
%   \item if $g \in G_R$ then $\Sigma(g) = R$,
%   \item if $g$ equals $\land_0$ or $\land_1$ then $\Sigma(g) = \land[n^k]$
%   \item if $g$ equals $\land_{\text{out}}$ then $\Sigma(g) = \land[2]$,
%   \item if $g \in G_\rank$ let $\Sigma(g) = \rank^r_p [a,b]$, and
%   \item if $g \in G_{\text{in}}$ and $g$ then $\Sigma(g) = \land[1]$.
%   \end{itemize}
%   Define $L$ as follows. For each $g \in G$,
%   \begin{itemize}
%     \setlength\itemsep{0mm}
%   \item if $g \in G$ and $g = g_{\text{out}}$ then for each $i \in [2]$,
%     $L(g)(i) := g_{\rank, i}$,
%   \item if $g \in G_\rank$ and $g = g_{\rank, i}$ then $L(g)(p,q) = g_{i,
%     (p,q)}$,
%   \item if $g \in G_{\land}$ and $g = g_{i, (p,q)}$ then $L(g)(1) = \land_1$
%     if $(p,q) \in E_i$ and $L(g)(1) = \neg_0$ otherwise,
%   \item if $g = \neg_0$ then $L(g)(1) = \land_0$, and
%   \item if $g = \land_0$ or $g = \land_1$ then for all $p \in [n^k]$ we have
%     that $L(g)(p)$ equals $\Lambda^{-1}(\vec{c}_p)$, where $\vec{c}_p$ is the
%     $p$th element of $[n]^k$ in the lexicographical ordering on $[n]^k$.
%   \end{itemize}

%   We note that for $i \in [2]$, $L(g_{\rank, i})(p,q) = \land_1$ if $(p,q)$ is
%   an edge in in $B_i$ and $L(g_\rank, i)(p,q) = \neg_0$ if $(p,q)$ is not an
%   edge in $B_i$. We thus have that $B_1$ and $B_2$ are bipartite isomorphic
%   (in the sense defined above) if, and only if, $g_{\rank, 1}$ is
%   syntactically equivalent to $g_{\rank, 2}$.

%   Since the construction of $C_n$ can be implemented in time polynomial in the
%   combined sizes of the input graphs, the mapping of $(B_1, B_2)$ to the tuple
%   $(C_n, (g_{\rank, 1}, g_{\rank, 2}))$ witnesses the reduction. The result
%   follows.
% \end{proof}
\begin{prop}
  Let $\tau$ be any non-empty relational vocabulary and let $n, r, p \in \nats$,
  with $p$ prime. There is a a polynomial time reduction from the graph
  isomorphism problem to the problem of deciding if a given pair of gates in a
  given symmetric rank-circuit for structures of size $n$ that (i) has injective
  labels, (ii) contains no constant gates, and (iii) contains at most two rank
  gates with bound $r$ and prime $p$, are syntactically equivalent.
  \label{prop:syntactic-graph-iso}
\end{prop}
\begin{proof}
  We have from \cite{} that there is a polynomial time reduction from the graph
  isomorphism problem to the bipartite isomorphism problem. The bipartite
  isomorphism problem is the decision problem for the set of all pairs of
  partitioned bipartite graphs $B_1 := (U_1, V_1, E_1)$ and $B_2 := (U_2, V_2,
  E_2)$ such that there exists a graph isomorphism $\pi : B_1 \rightarrow B_2$
  with $\restr{\pi}{U_1} = U_2$ and $\restr{\pi}{V_1} = V_2$.

  Suppose we are given two partitioned bipartite graphs $B_1 := (U_1, V_1, E_1)$
  and $B_2 := (U_2, V_2, E_2)$. We assume, without a loss of generality, that
  there exists $a, b \in \nats$ such that $U_1 = U_2 = [a]$ and $V_1 = V_2 =
  [b]$.
